### 2013-02-23 article/paper.tex
added Neo4J implementation details, and data-sets information

\subsection{Redis}
Redis is an open-source, in-memory key-value store database system, sponsored
by VMware. It's written in C and has atomic operations, we used the following features:
\begin{itemize}
	\item {\bf Transactions}
        You can use redis transaction to save time spent on network trafic 
        between the client API and the server. We used that feature to speed up
        the graph load significantly.
	\item {\bf Data structure as value}
        Redis support storing many types of data structures as valuse and 
        performe efficient, atomic operation over them. I.e. there are strings,
        hashes (key-value pairs), lists, sets (in which we used to store the 
        adjacency lists), and sorted-sets (like sets, but with rank for each item).
        Operations include: incrementing value in hash, appending to lists, 
        union/intersection/difference on sets, and getting sorted-set member with top rank.
	\item {\bf Lua Scripting}
        Lua scripting enable you to run your own code, in an atomic matter inside the server.
        Unlike transactions, you can add non-Redis logic to this atomic operation. We used 
        Lua scripting to speed-up and algorithem which took long time to run. \footnote{Sara: 
        we are kind of abusing this feature, by locking the db and assuming single client, 
        should we mention that?}
\end{itemize}

@@

\subsection{Neo4J}
Neo4J is a multi-platform, open-source, graph database supported by Neo
Technology. Some of it's main features are:
\begin{itemize}
    \item {\bf transactions}
        Neo4J supports ACID transactions, similar to RDBMS. This also enables 
        you to send chunk of commands to the database instead of single 
        instactions and avoid network latency for each call. We tried to use
        this to load our graph, but this method is just too slow for bulk insertions.
    \item {\bf Embedded mode}
        Embedding Neo4J in your application, mainly saves time on REST calls 
        and enables fast interaction with the database. We used that feature 
        to load the graph, and the move the generated files to the server.
        We found that currently, this is the fastest way to bulk insert a 
        large graph to Neo4J.
    \item {\bf Cypher query language}
        Cypher is a declerative query language for graphs, enabling you to
        match pattern, filter on properties and make changes. It's syntax is
        much like SQL, e.g. friends of friends which are not directly connected:
        \begin{verbatim}
            START n=node(...)
            MATCH n-->m-->o
            WHERE not ( n-->o )
            RETURN o
        \end{verbatim}
    \item {\bf Lucene Indexing}
\end{itemize}
ACID transactions, expressive graph query
language, indexing and high availability. It can operate in an embedded or
client-server architecture.

In this database, we attempted to use only Neo4j's declarative query language
(Cypher) to implement the algorithms, and the client-server architecture to run
them with. The benchmarking was performed using the Python API, which also
assisted in implementing some multi-query algorithms. We found that the best
way to load the graph into Neo4J was to use the BatchInserter class in the
embedded architecture, and copy the created files into the server.

There are some limitations to the Cypher language users should be aware of:
\begin{itemize}
\item There are not many built in functions (e.g. no Power, Log, or Rand
  functions) 
\item No support for user defined functions
\item Queries are really slow when matching a non-trivial pattern over the whole
graph (e.g. common neighbors)  
\end{itemize}

\section{Experiments}
We benchmarked the above implementations using 10 undirected graphs:
\begin{itemize}
\item Three graphs were extracted from the DBLP XML records using the authors as
nodes, and co-authorship relation as edges. Each graph was over a different
time period, and  we only kept authors with 3 or more publication within the
selected time period (core-3).

\item Seven graphs from the Stanford Network Analysis Project (SNAP) dataset
  collection. No changes were made to these graphs.
\end{itemize}

>>>>

\subsection{Redis}
Redis is an open-source, in-memory key-value store database system, sponsored
by VMware. It's written in C and has atomic operations, we used the following features:
\begin{itemize}
	\item {\bf Transactions}
        You can use redis transaction to save time spent on network trafic 
        between the client API and the server. We used that feature to speed up
        the graph load significantly.
	\item {\bf Data structure as value}
        Redis support storing many types of data structures as valuse and 
        performe efficient, atomic operation over them. I.e. there are strings,
        hashes (key-value pairs), lists, sets (in which we used to store the 
        adjacency lists), and sorted-sets (like sets, but with rank for each item).
        Operations include: incrementing value in hash, appending to lists, 
        union/intersection/difference on sets, and getting sorted-set member with top rank.
	\item {\bf Lua Scripting}
        Lua scripting enable you to run your own code, in an atomic matter inside the server.
        Unlike transactions, you can add non-Redis logic to this atomic operation. We used 
        Lua scripting to speed-up and algorithm which took long time to run. \footnote{Sara: 
        we are kind of abusing this feature, by locking the db and assuming single client, 
        should we mention that?}
\end{itemize}

@@

\subsection{Neo4J}
Neo4J is a multi-platform, open-source, graph database supported by Neo
Technology. Some of it's main features are:
\begin{itemize}
    \item {\bf Transactions}
        Neo4J supports ACID transactions, similar to RDBMS. This also enables 
        you to send chunk of commands to the database instead of single 
        instactions and avoid network latency for each call. We tried to use
        this to load our graph, but this method is just too slow for bulk insertions.
    \item {\bf Embedded mode}
        Embedding Neo4J in your application mainly saves time on REST calls 
        and enables fast interaction with the database. We used that feature 
        to load the graph, and then move the generated files to the server.
        We found that currently, this is the fastest way to bulk insert a 
        large graph into a Neo4J server.
    \item {\bf Cypher query language}
        Cypher is a declerative query language for graphs, enabling you to
        match pattern, filter on properties and make changes. It's syntax is
        much like SQL, e.g. friends of friends which are not directly connected:
        \begin{verbatim}
            START n=node(...)
            MATCH n-->m-->o
            WHERE not ( n-->o )
            RETURN o
        \end{verbatim}
    \item {\bf Indexing}
		There are two main indexing options in Neo4J:
		\begin{enumerate}
			\item {\bf Using the graph}
				An index can be implemented as a layed over the graph, 
				i.e. by adding more relations which represent the index
				structure.
			\item {\bf Separate index}
				By default, Neo4J uses Apache Lucene as a backend index engine.
				This enables you to index Nodes and Relations, and use all Lucene's
				features, including fulltext index.
				We used this form of indexing to speed-up the Preferential attachment
				algorithm.
		\end{enumerate}
\end{itemize}

We attempted to implement all the algorithms using Cypher. Though this was 
usually an easy task, some of the language current limitations prevented us 
from achiving that goal:
\begin{itemize}
	\item Lack of major math functions (e.g. Power, Log, Rand), and no support 
		for user defined function, prevented from implementing the Adamic/Adar 
		algorithm or to use multiple queries, joined by logic in Python, 
		to implement the Katz algorithm.
	\item Performance issues on mutiple start nodes, in particular when matching
		a non-trivial pattern over the whole graph (i.e. our global benchmarks).
		We had to omit the global version of Jaccard Coefficient, and find another
		approach to the global Common Neighbors implementation.
\end{itemize}

Implementation details:
\begin{itemize}
	\item {\bf Helper Neighbor count property}
		Since Neo4J doen't provide a fast way to get the number of node's neighbors,
		we added a neighbor count property to each node. This was useful for 
		almost any algorithm we implemented.
	\item {\bf Helper Index}
		We generated an Lucene index containing 101 nodes with most neighbors,
		to speed up the Preferential attachment algorithm.
	\item {\bf Common Neighbors (global)} 
		In order to circumvent a Cypher performance issue with multiple start 
		nodes, we used a different approach to the Common Neighbors implementation:
		we realized that the result-set of such query will only contain nodes with
		at least $\alpha$ number of neighbors, where $\alpha$ is number of neighbors
		of the lowers scoring pair in the result-set.
		So we used a binary search method to find that $\alpha$, by iteratively 
		running the	query over nodes with $\beta$ number of neighbors, where $\beta$
		is initiated to $M/2$ ($M$ is the maximum number of neighbors a node in the
		graph has), divding $\beta$ by two each iteration, until $\alpha > \beta$.
	\item {\bf Preferential attachment (Global)}
        We used the helper index that we created beforehand, to find pairs with 
		maximal product.
	\item {\bf Common Neighbors (for node)}
        We used a simple query.
	\item {\bf Jaccard's Coefficient (for node)}
        We used a simple query, using the relation between the cardinality of 
		union and intersection
	\item {\bf Preferential attachment (for node)}
        We used the helper index that we created beforehand, to find pairs with 
		maximal product.
	\item {\bf Graph Distance (for node)}
		We iteratively ran a query, returning nodes until depth $i$ (increased on 
		every iteration), until depth limit or returend data set limit has met.
	\item {\bf Katz (for node)}
		Without power function support in Cypher, we had to use multiple queries,
		to calculate the Katz algorithm.
	\item {\bf Rooted PageRank (for node)}
        We used an iterative loop with the same logic like the other two databases,
		storing the intermediate values as node properties.
\end{itemize}

\section{Experiments}
We benchmarked the above implementations using 10 undirected graphs:
\begin{itemize}
	\item 
		Three graphs were extracted from the DBLP XML records using the authors 
		as nodes, and co-authorship relation as edges. Each graph was over a 
		different time period, and  we only kept authors with 3 or more 
		publication within the selected time period (core-3). 
	\item 
		Seven graphs from the Stanford Network Analysis Project (SNAP) dataset
		collection:
		\begin{enumerate}
			\item {ego-Facebook} 
				Anonymized Facebook data that was collected from survey 
				participants using a Facebook app.
			\item {email-Enron} 
				Enron email communication network, released by the 
				Federal Energy Regulatory Commission during its investigation.
			\item {ca-AstroPh}
				Scientific collaborations between authors papers submitted to 
				Astro Physics category.
			\item {ca-CondMat}
				Scientific collaborations between authors papers submitted to 
				Condense Matter category.
			\item {ca-GrQc}
				Scientific collaborations between authors papers submitted to 
				General Relativity and Quantum Cosmology category.
			\item {ca-HepPh}
				Scientific collaborations between authors papers submitted to 
				High Energy Physics - Phenomenology category.
			\item {ca-HepTh}
				Scientific collaborations between authors papers submitted to 
				High Energy Physics - Theory category
		\end{enumerate}
\end{itemize}

\begin{tabular}{ |l||c|r| }
	\hline
	Data-set & Node Count & Edge Count \\
	\hline
	dblp-2002-2009-core3 & 182493 & 1621846 \\
	dblp-2010-2012-core3 & 248695 & 2589320 \\
	dblp-all-core3       & 366600 & 4349796 \\
	ego-Facebook         & 4039   & 170174  \\
	email-Enron          & 36692  & 367662  \\
	ca-AstroPh           & 18771  & 396160  \\
	ca-CondMat           & 23133  & 186936  \\
	ca-GrQc              & 5241   & 28980   \\
	ca-HepPh             & 12006  & 237010  \\
	ca-HepTh             & 9875   & 51971   \\
	\hline  
\end{tabular}
